Models of gravitational lens candidates from Space Warps CFHTLS

### 2017-06-15 parts/morph.tex
editing

The fourth diagnostic is based on the fact that the arrangement of
lensed images of a pointlike source through a non-circular
gravitational lens depends on the location of the source relative to
the long and short axes of the lens.  This dependence is quite robust
and independent of many other details of the lensing mass
distribution.  Sources close to being dead-centre behind a lens tend
to produce quads, sources at larger transverse distance tend to
produce doubles.  We denote these as Q and D respectively, and add a
prefix to the Q systems, as follows: we write LQ if the source is
inferred as displaced along the long axis of the lens, SQ if displaced
along the short axis, IQ if inclined to both axes, CQ if only very little to
no inclination.  Although the
unlensed source and its location are obviously not seen, the LQ, SQ, IQ and CQ
cases correspond to easily-seen image morphologies \citep[see,
  e.g.,][]{2003AJ....125.2769S}.
\begin{itemize}
\item The simplest is the SQ case: this creates a saddle point and two
  minima in an arc, with the second saddle point on the other side of
  the lensing galaxy, and closer to the galaxy than the arc is.  SW58
  and SW28 in the upper row of Figure~\ref{fig:splinput} are typical
  examples of SQ.
\item The middle row of Figure~\ref{fig:splinput} shows three IQ
  systems, SW05, SW42 and SW19.  This type has a characteristic
  asymmetry, often with two images close together.
\item The lower row of Figure~\ref{fig:splinput} shows three LQ
  systems, SW09, SW29 and SW02, and the failed model SW57 at top right
  also possibly belonging to this category.  Here the images have a
  fairly symmetric arrangement with an arc and a counter-image on the
  other side, but the spaghetti diagram is completely asymmetric.  LQ
  can be visually distinguished from SQ from the relative distances of
  the arc and the counter-image.  For LQ the arc is closer, for SQ the
  counter-image is closer.
\end{itemize}
Although the labels `short-axis quad' and so on are not standard in
the literature, the morphological classification they express is
familiar to experienced modellers.  Hence they can be useful to
researchers wishing to apply other modelling methods to the same
systems.

>>>>

The fourth diagnostic is based on the fact that the arrangement of
lensed images of a pointlike source through a non-circular
gravitational lens depends on the location of the source relative to
the long and short axes of the lens.  This dependence is quite robust
and independent of many other details of the lensing mass
distribution.  Sources close to being dead-centre behind a lens tend
to produce quads, sources at larger transverse distance tend to
produce doubles.  We denote these as Q and D respectively, and add a
prefix to the Q systems, as follows: we write LQ if the source is
inferred as displaced along the long axis of the lens, SQ if displaced
along the short axis, IQ if inclined to both axes, CQ if only very little to
no displacement.  Although the
unlensed source and its location are obviously not seen, the LQ, SQ, IQ and CQ
cases correspond to easily-seen image morphologies \citep[see,
  e.g.,][]{2003AJ....125.2769S}.
\begin{itemize}
\item The simplest is the SQ case: this creates a saddle point and two
  minima in an arc, with the second saddle point on the other side of
  the lensing galaxy, and closer to the galaxy than the arc is.  SW58
  and SW28 in the upper row of Figure~\ref{fig:splinput} are typical
  examples of SQ.  If the source would move outwards along the short
  axis, the minimum-saddle-minimum would merge into a single minimum,
  leaving a D system; the transition is known as cusp catastrophe.
\item The middle row of Figure~\ref{fig:splinput} shows three IQ
  systems, SW05, SW42 and SW19.  This type has a characteristic
  asymmetry, often with two images close together.  If the source
  would move outwards, the minimum-saddle pair would merge and
  mutually cancel, again leaving a D system; the transition is known
  as a fold catastrophe.
\item The lower row of Figure~\ref{fig:splinput} shows three LQ
  systems, SW09, SW29 and SW02, and the failed model SW57 at top right
  also possibly belonging to this category.  Here the images have a
  fairly symmetric arrangement with an arc and a counter-image on the
  other side, but the spaghetti diagram is completely asymmetric.  If
  the source would move outwards along the long axis, the
  saddle-minimum-saddle would merge into a single saddle --- another
  form of cusp catastrophe.  LQ can be visually distinguished from SQ
  from the relative distances of the arc and the counter-image.  For
  LQ the arc is closer, for SQ the counter-image is closer.
\end{itemize}
CQ systems are often called `cross' or `Einstein-cross' systems; IQ
systems are sometimes called `folds', with `cusp' commonly used for
both SQ and LQ.  The labels `short-axis quad' and so on are not
standard in the literature, but the morphological classification they
express is familiar to experienced modellers.  Hence they can be
useful to researchers wishing to apply other modelling methods to the
same systems.
